Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-17 manuscript/manuscript.Rmd
more pics and fix refs

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
---

@@

(TODO: link to github repo)

@@

These model architectures are relatively uncomplicated by today's standards, and are so chosen to better highlight the
embedding components. Later, in Section X (TODO), we will investigate more involved architectures utilizing embeddings
that represent the state of the art for modeling tabular data.

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). 
The categorical inputs go through one-dimensional embedding
layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which
have been normalized in data pre-processing, before being sent to a hidden linear layer of 64 units. A ReLU (TODO: reference an equation in prev. section) activation
is then applied to the outputs before they are sent to a final output layer resulting in a scalar. To constrain
predictions to be positive, we apply softplus activation, which is defined to be

$$
x \mapsto \log(1 + e^x). 
$$

@@

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings
for the `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values
to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think
of the embedding values for each variable as a continuous represention of the variable itself, rather than a 
"coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a
table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities
within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable,
which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table 
\@ref(tab:relativities). In Section TODO, we will discuss techniques to visualize these embeddings.

@@

```{r num-floors, echo=FALSE, fig.cap="Number of floors", out.width = "90%", fig.align="center"}
knitr::include_graphics("images/num_floors.png")
```

```{r flood-zone, echo = FALSE, fig.cap="Flood zone embeddings", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/flood_zone.png")
```

@@

@@

## Model Training and Results

@@

# Attention based modelling

@@

>>>>

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.pos = "!h", out.extra = "")
```

@@

@@

These model architectures are relatively uncomplicated by today's standards, and are so chosen to better highlight the
embedding components. Later, in Section \ref{attn}, we will investigate more involved architectures utilizing embeddings
that represent the state of the art for modeling tabular data.

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). 
The categorical inputs go through one-dimensional embedding
layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which
have been normalized in data pre-processing, before being sent to a hidden linear layer of 64 units. A ReLU activation,

\begin{equation}
x \mapsto \max(0, x),
\end{equation}

is then applied to the outputs before they are sent to a final output layer resulting in a scalar. To constrain
predictions to be positive, we apply softplus activation, which is defined to be

\begin{equation}
x \mapsto \log(1 + e^x). 
\end{equation}

@@

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings
for the `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values
to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think
of the embedding values for each variable as a continuous represention of the variable itself, rather than a 
"coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a
table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities
within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable,
which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table 
\@ref(tab:relativities). In Section \ref{interpretation}, we will discuss techniques to visualize these embeddings.

@@

```{r num-floors, echo=FALSE, fig.cap="Number of floors", out.width = "80%", fig.align="center"}
knitr::include_graphics("images/num_floors.png")
```

```{r flood-zone, echo = FALSE, fig.cap="Flood zone embeddings", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone.png")
```

@@

(TODO: insert brief theoretical stuff). See Figure \@ref(fig:flood-zone-pca).

```{r flood-zone-pca, echo = FALSE, fig.cap="Flood zone PCA", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone_pca.png")
```

@@

(TODO: insert brief theoretical stuff). See Figure \@ref(fig:flood-zone-tsne).

```{r flood-zone-tsne, echo = FALSE, fig.cap="Flood zone t-SNE", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone_tsne.png")
```

## Training and Inference

@@

# Attention based modelling {#attn}

@@

\newpage

# References {.unnumbered}
